---
title: "p8105_hw3_rj2543"
author: "rj2543"
date: "October 10, 2018"
output: github_document
---

```{r set up, include = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

# Problem 1

This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the  p8105.datasets package.

First, do some data cleaning:

* format the data to use appropriate variable names;

* focus on the “Overall Health” topic

* include only responses from “Excellent” to “Poor”

* organize responses as a factor taking levels ordered from “Excellent” to “Poor”

```{r problem 1 data cleaning}
data("brfss_smart2010")

brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% # clean up variable names
  filter(topic == "Overall Health") %>% # focus on "Overall Health" topic
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%  # include responses from "Excellent" to "Poor"
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"), ordered = TRUE)) # "response" as factor taking levels ordered
  
brfss
```

Using this dataset, do or answer the following (commenting on the results of each):

* In 2002, which states were observed at 7 locations?

```{r answer 1}
brfss %>% 
  filter(year == 2002) %>% # fix "year" to 2002
  distinct(locationabbr, locationdesc) %>% # "distinct" locations
  count(locationabbr) %>% # count observations
  filter(n == 7)
```

```{r answer 1 another approach}
brfss %>% 
  filter(year == 2002) %>% 
  group_by(locationabbr) %>% 
  summarise(n = n_distinct(locationdesc)) %>% 
  filter(n == 7)
```

**In 2002, three states were observed at 7 locations, which are CT, FL and NC in specific.**

* Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r spaghetti plot}
brfss %>% 
  group_by(year, locationabbr) %>% 
  summarise(n = n_distinct(locationdesc))%>% 
  ggplot(aes(x = locationabbr, y = n, color = year)) + 
  geom_line()
  
```


* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

* For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time


